# Logging and Monitoring

**Learning objectives:**

-   Differentiate between logs and metrics.
-   Log events from your code using the {log4r} package.
-   Recognize some open source options for emitting metrics.

## Logs and Metrics {-}

-   Logs = events as they happen.
-   Metrics = measurements of how things are working.

## {log4r} Basics {-}

-   {log4r} vs {base} vs {cli}
    -   {base} message/warning/error have least structure.
    -   {cli} adds structure for human readability.
    -   {log4r} adds structure that's useful for a stream of events (timestamp, etc)
-   The object created via `logger()` defines where things go & how they look.

```{r 04-log4r defaults, eval = FALSE}
library(log4r)
log <- logger(threshold = "INFO", appenders = console_appender()) # Defaults

info(log, "Log starting.")
warn(log, "This is a warning!")
error(log, "This is an error, but not an R error.")
debug(log, "This won't show up.")
```

```{r 04-logr result, eval = FALSE}
## INFO  [2022-12-03 10:31:41] Log starting.
## WARN  [2022-12-03 10:31:41] This is a warning!
## ERROR [2022-12-03 10:31:41] This is an error, but not an R error.
```

## {log4r} Layouts {-}

-   Alex hinted at this but didn't dig into it.

```{r 04-logr json, eval = FALSE}
log <- logger(appenders = console_appender(layout = json_log_layout()))

info(log, "Log starting.")
warn(log, "This is a warning!")
error(log, "This is an error.")
debug(log, "This won't show up.")
## {"message":"Log starting.","level":"INFO","time":"2022-12-03T16:35:23Z"}
## {"message":"This is a warning!","level":"WARN","time":"2022-12-03T16:35:23Z"}
## {"message":"This is an error.","level":"ERROR","time":"2022-12-03T16:35:23Z"}
```

```{r 04-logr fancier json, eval = FALSE}
info(log, message = "Log starting.", my_var = 2)
## {"message":"Log starting.","my_var":2,"level":"INFO","time":"2022-12-03T16:37:19Z"}
```

```{r 04-logr logfmt, eval = FALSE}
log <- logger(appenders = console_appender(layout = logfmt_log_layout()))
info(log, message = "Log starting.", my_var = 2)
## level=INFO ts=2022-12-03T16:40:13Z message="Log starting." my_var=2
```

## Metrics {-}

-   [Prometheus](https://prometheus.io/) = open source monitoring tool
-   [Grafana](https://grafana.com/) = open source viz for Prometheus
-   [{openmetrics}](https://github.com/atheriel/openmetrics) = R package to add metrics to {plumber} & {shiny}
-   [{vetiver}](https://rstudio.github.io/vetiver-r/) = RStudio's model deployment framework, includes metrics
-   Hoping to revisit this in January ("What has changed?" meeting)

## Review Discussion {-}

-   Differentiate between logs and metrics.
-   Log events from your code using the {log4r} package.
-   Recognize some open source options for emitting metrics.


## Meeting Videos {-}

### Cohort 1 {-}

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
